---
title: "a3_task4_nicholas_bissonnette"
author: "Nichoals Bissonnette"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message= FALSE, warning= FALSE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

#### Task 4 

Step 3. I will read in the urchins data in the first code chunk, I will clean the dates and create new column titles as well for month and year. 
```{r}
urchins <- read.csv(here("data", "urchins.csv")) %>% 
  clean_names()

urchins_clean <-urchins %>% 
  mutate(Date=mdy(date)) %>% 
  mutate(month = month(Date)) %>% 
  mutate(year= year(Date))


```

a. Now I will wrangle the data to only include Naples Reef observations for purple urchins observed in January 2010 using the filter().

```{r}
purple_jan_10 <- urchins_clean %>% 
  filter(common_name == "Purple Urchin", year == 2010, month == 1, site== "NAPL")
```

b. Now I will make histograms to explore the size distribution for each of the treatments. 

```{r}
ggplot(data= purple_jan_10, aes(x= size)) +
  geom_histogram(bins=15)+
  facet_wrap(~treatment)
```
In the next chunk I will create a quantile-quantile plot to assess normality. 

```{r}
ggplot(data= purple_jan_10, aes(sample= size)) +
  geom_qq()+
  facet_wrap(~treatment)
```

The above qqplot shows a linear relationship in the distribution indicating normality. The histograms show a slightly negative skew for the annual treatment group, but overall both distributions show normality.  

C. Now I will create a table of summary statistics including mean, standard deviation, median and sample size.
```{r}
purple_stats<- purple_jan_10 %>% 
  group_by(treatment) %>% 
  summarize(mean_size = mean(size),
            sd_size = sd(size),
            median_size= median(size),
            n= n())
purple_stats %>% 
  kbl() %>% 
  kable_styling()
```

d. Now I will preform a two-sample t-test to compare means between two groups. 
```{r}
annual_sample <- purple_jan_10%>% 
  filter(treatment== "ANNUAL") %>% 
pull(size)

control_sample <- purple_jan_10 %>% 
  filter(treatment== "CONTROL") %>% 
  pull(size)

t.test(x= annual_sample, control_sample)
```
e-f. The T-test shows that the annual treatment group has a higher mean value of 3.92  then the control group which is 3.73. The numerical difference in means is  3.92- 3.37 = 0.55. For the collected samples the annual treatment urchins have a larger size of 0.55 cm. The p-value tells us there is about a 25% chance that both of these samples came from populations with the same mean. 
