---
title: "a3_task3_nicholas_bissonnette"
author: "Nichoals Bissonnette"
date: "11/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

#### Section 1
Read in the urchins data  to compare distribution sizes. 

```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()

```

In this code chunk I will create a histogram of the sizes of both red and purple urchins.

```{r}
ggplot(data= urchins, aes(x= size)) +
  geom_histogram(bins= 13)+
  facet_wrap(~common_name)



```

In this code chunk I will create a box plot.

```{r}
ggplot(data= urchins, aes(x= common_name, y= size)) +
  geom_boxplot()
```

ii. Now I will make a descriptive statistics table using group_by(common_name) and summarize() and making it more presentable using kabelExtra.

```{r}
summary_stats <- urchins %>% 
  group_by(common_name) %>% 
  summarize(
    mean_size= mean(size),
    meidan_size = median(size),
    sd_size = sd(size),
    Sample_count = n()
  )
summary_stats %>% 
  kbl() %>% 
  kable_styling()
  
```

iii. After looking at the exploratory data and tables, I observe that the both purple and red urchins have a normal distribution and the purple urchin has smaller data spread when compared to red. The difference between the purple urchin median and mean values is about %5 while the difference in red median and mean values are about %3. 

#### Section 2

```{r}
urchins_clean <-urchins %>% 
  mutate(date=mdy(date))
class(urchins_clean$date)
```

i. I will now create a new subset of data containing only purple urchins data points. Also I will convert the date data to the "Date" class, also creating a new column that only contains the year. Lastly I will create a new column that only has full site name. But I also need to clean the dates to get them in the proper format. 
```{r}


purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(Date= ymd(date)) %>% 
  mutate(year= year(Date)) %>% 
  mutate(site_full= case_when(
     site %in% c("NAPL") ~ "Naples Reef",
     site %in% c("AQUE") ~ "Arroyo Quemado Reef",
     site %in% c("CARP") ~ "Carpinteria Reef",
     site %in% c("MOHK") ~ "Mohawk Reef",
     site %in% c("IVEE") ~ "Isla Vista Reef"
    
  ))
```

ii. Next I will create a summary table containing the total number of purple urchins observed by site and year using group_by() and count().

```{r}
urchins_by_site_and_year <- purple_urchins %>%
  group_by(site_full, year) %>% 
    count()
urchins_by_site_and_year
```

iii. Now I will create an exploratory graph of the counts of purple urchins recorded at each site over time for the years.

```{r}
ggplot(data= urchins_by_site_and_year, 
       aes(x=year, y= n))+
  geom_line( aes(color= site_full))
```

